---
output: rmarkdown::github_document
always_allow_html: yes
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
```

# STAT 585 Lab 4 - Group 5
 
**Katherine Goode, Gina Nichols, Xiyuan Sun, Ying Zheng**  
**04/21/2019**

This document contains an overview of our work for lab 4. Our GitHub repository for this lab can be found [here](https://github.com/xiyuansun/stat585Lab4).

```{r message = FALSE}
# Libraries used
library(lubridate) # for working with dates
library(zoo)
library(ggplot2)  # for creating graphs
library(scales)   # to access breaks/formatting functions
library(gridExtra) # for arranging plots
library(grid)   # for arrangeing plots
library(dplyr)  # for subsetting by season
library(tidyverse)
```

## Iowa Liquor Sales Data

#### Demonstration of Accessing the Data from the API

The code below demonstrates how to access the data from the API. Since we do not have a key, this accesses a subset of the data.

```{r}
story_liquor_raw_subset <- jsonlite::fromJSON("https://data.iowa.gov/resource/m3tr-qhgy.json")
```

Below is a glimpse of the obtained dataset.

```{r}
glimpse(story_liquor_raw_subset)
```

#### Data Cleaning

For the Shiny app, we will use the full Story County liquor sales dataset provided by Dr. Hofmann. This data is read in below.

```{r}
# Read in data
story_liquor_raw <- read_csv("data/Iowa_Liquor_Sales-Story.csv", col_types = cols())
```

We cleaned up the data before using it in the Shiny app. This involved renaming the variables, adjusting the variable types, and dealing with typos. This work is all included in the following section of code.

```{r dat clean}
# Clean data
story_liquor_cleaned <- story_liquor_raw %>%
  
  # Rename cols
  rename_all(funs(
  str_replace_all(., "[\\(\\)_\\/ ]", "") %>%
  tolower())) %>%
  
  # Adjust date variable
  mutate(date = lubridate::mdy(date)) %>%

  # Send character variables to lower
  mutate_if(.predicate = is.character,
            .funs = tolower) %>%
  
  # Create factor variables where necessary
  mutate_at(.vars = c("storenumber", 
                      "storename", 
                      "address", 
                      "city",
                      "zipcode", 
                      "storelocation", 
                      "countynumber",
                      "county",
                      "category", 
                      "categoryname", 
                      "vendornumber", 
                      "vendorname", 
                      "itemnumber", 
                      "pack"),
            .funs = factor) %>%
  
  # Clean up factor variables
  mutate(address = str_replace(address, ",", ""),
         address = str_replace(address, "issac", "isaac"),
         address = str_replace(address, "4518 mortenson rd ste 109", "4518 mortonsen street suite #109"),
         address = str_replace(address, "lincolnway", "lincoln way"),
         storename = str_replace(storename, "a j's", "aj's"),
         storename = str_replace(storename, "casey's general store # 2560/ ames", 
                                 "casey's general store # 2560"),
         storename = str_replace(storename, "casey's general store # 2560", 
                                 "casey's general store #2560 / ames"), 
         storelocation = str_split(storelocation, "\n", simplify = TRUE)[,3],
         latitude = str_split(storelocation, ", ", simplify = TRUE)[,1] %>% 
           str_replace("\\(", "") %>%
           as.numeric(),
         longitude = str_split(storelocation, ", ", simplify = TRUE)[,2] %>%
           str_replace("[\\)]", "") %>%
           as.numeric(),
         categoryname = str_replace(categoryname, "american cordials & liqueurs", 
                                    "american cordials & liqueur"),
         categoryname = str_replace(categoryname, "american distilled spirits specialty", 
                                    "american distilled spirit specialty"),
         categoryname = str_replace(categoryname, "american vodkas", "american vodka"),
         categoryname = str_replace(categoryname, "cocktails /rtd", "cocktails / rtd"),
         categoryname = str_replace(categoryname, "flavored gins", "flavored gin"),
         categoryname = str_replace(categoryname, "imported distilled spirits specialty", 
                                    "imported distilled spirit specialty"),
         categoryname = str_replace(categoryname, "imported vodkas", "imported vodka"),
         vendorname = str_replace_all(vendorname, "[,\\.\\/-]", "")) %>%
  group_by(vendornumber) %>%
  mutate(vendorname = vendorname[1]) %>%
  ungroup() %>%
  mutate(vendorname = factor(vendorname)) %>%
  
  # Get rid of weird extremes
  filter(saledollars > 0,
  volumesoldliters > 0,
  volumesoldgallons > 0,
  volumesoldliters < 250) %>%
  
  # Assign new simple categories
  mutate(
    catsimp = factor(case_when(
      grepl("amaretto", categoryname) ~ "amaretto",
      grepl("brand", categoryname) ~ "brandy",
      grepl("creme", categoryname) ~ "creme",
      grepl("cocktail|sec", categoryname) ~ "general_mixer",
      grepl("spirit", categoryname) ~ "general_spirit",
      grepl("gin", categoryname) ~ "gin",
      grepl("liqueur", categoryname) ~ "liqueur",
      grepl("schnap", categoryname) ~ "schnapps",
      grepl("rum", categoryname) ~ "rum",
      grepl("special", categoryname) ~ "specialty",
      grepl("tequ|mez", categoryname) ~ "tequila",
      grepl("vodka", categoryname) ~ "vodka",
      grepl("whisk|bourb|scotch|rye", categoryname) ~ "whisky"
      ))
    ) %>%
  
  # Select the final order of the variables
  select(invoiceitemnumber:storelocation, 
         latitude, 
         longitude, 
         countynumber:categoryname, 
         catsimp, 
         vendornumber:volumesoldgallons)
```

A glimpse of the cleaned data is shown below.

```{r}
# Glimpse the data
glimpse(story_liquor_cleaned)
```

```{r eval = FALSE, echo = FALSE}
# Save the cleaned data
saveRDS(story_liquor_cleaned, "./data/story_liquor_cleaned.rds")
```

## Shiny App Visualizations

#### Preparing Temporal Data

For the temporal visualizations, we used the cleaned data to compute the number of sales and average cost per liter (in dollars per liter) by liquor category for each date in the data set.

```{r}
# Prepare temporal data

story_temporal_data <- story_liquor_cleaned %>%
  mutate(dollar_per_liter =saledollars/volumesoldliters)%>%
  select(date, storename, latitude, longitude, catsimp, vendorname,dollar_per_liter)%>%
  na.omit()

story_temporal_data$month_yr <- format(as.Date(story_temporal_data$date), "%Y-%m")
story_temporal_data$yr <- year(story_temporal_data$date)
story_temporal_data <- story_temporal_data[order(as.Date(story_temporal_data$date)),]

```

```{r eval = FALSE, echo = FALSE}
# Save the spatial data
saveRDS(story_temporal_data, "./data/story_temporal_data.rds")
```



#### Preparing Spatial Data

For the spatial visualizations, we used the cleaned data to compute the number of sales and average cost per liter (in dollars) by liquor category for each store in the data set.

```{r}
# Prepare spatial data
story_spatial_data <- story_liquor_cleaned %>%
  
  # Select desired variables and remove any na's
  select(storename, latitude, longitude, catsimp, saledollars, volumesoldliters) %>%
  na.omit() %>%
  
  # Compute dollar per liter for each observation 
  mutate(dollar_per_liter = saledollars / volumesoldliters) %>%
  
  # Round latitude and longitude to deal with differences in locations for the
  # same store
  mutate_at(.vars = c("latitude", "longitude"),
            .funs = round,
            digits = 1) %>%
  
  # Determine the number of sales and average dollar per liter
  group_by(storename, catsimp, latitude, longitude) %>%
  summarise(count = n(), 
            ave_dollar_per_liter = round(mean(dollar_per_liter), 2)) %>%
  ungroup()
```

```{r eval = FALSE, echo = FALSE}
# Save the spatial data
saveRDS(story_spatial_data, "./data/story_spatial_data.rds")
```
